---
title: "Supply Chain Resilience Network Analysis"
subtitle: "Interconnections and Correlations Between Dimensions"
format:
  pdf:
    documentclass: article
    geometry: margin=0.75in
    fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(igraph)
library(ggraph)
library(corrplot)

# Load data - find the project root dynamically
script_dir <- if (exists("knitr")) dirname(knitr::current_input()) else getwd()
project_root <- if (basename(script_dir) == "report_variations") dirname(script_dir) else script_dir
data_path <- file.path(project_root, "data", "cleaned_master.csv")
df <- read.csv(data_path, stringsAsFactors = FALSE)
```

# Executive Summary

This network analysis reveals the hidden relationships between resilience dimensions. Understanding these correlations helps identify which capabilities reinforce each other and which operate independently, enabling smarter investment decisions.

## Analysis Foundation

```{r foundation}
total_companies <- length(unique(df$company_name[df$company_name != "" & !is.na(df$company_name)]))
total_respondents <- nrow(df)

cat(sprintf("**Organizations Analyzed:** %d\n\n", total_companies))
cat(sprintf("**Data Points:** %d responses\n\n", total_respondents))
cat("**Network Principle:** Strong correlations indicate dimensions that typically develop together.\n")
```

\newpage

## Dimension Correlation Network

This network diagram visualizes relationships between the five resilience dimensions. Thicker edges indicate stronger correlations, while node positioning reflects relationship patterns.

```{r correlation-network, fig.width=12, fig.height=12}
# Calculate correlation matrix for dimensions
dimension_scores <- df %>%
  transmute(
    Redundancy = (up__r + in__r + do__r) / 3,
    Collaboration = (up__c + in__c + do__c) / 3,
    Flexibility = (up__f + in__f + do__f) / 3,
    Visibility = (up__v + in__v + do__v) / 3,
    Agility = (up__a + in__a + do__a) / 3
  )

# Calculate correlations
cor_matrix <- cor(dimension_scores, use = "pairwise.complete.obs")

# Create network from correlation matrix
# Only include correlations above threshold
threshold <- 0.3
edges <- data.frame()
for(i in 1:(nrow(cor_matrix)-1)) {
  for(j in (i+1):ncol(cor_matrix)) {
    if(abs(cor_matrix[i,j]) >= threshold) {
      edges <- rbind(edges, data.frame(
        from = rownames(cor_matrix)[i],
        to = colnames(cor_matrix)[j],
        weight = cor_matrix[i,j],
        abs_weight = abs(cor_matrix[i,j])
      ))
    }
  }
}

# Create graph
if(nrow(edges) > 0) {
  g <- graph_from_data_frame(edges, directed = FALSE)

  # Calculate node metrics
  V(g)$degree <- degree(g)
  V(g)$avg_score <- sapply(V(g)$name, function(dim) {
    mean(dimension_scores[[dim]], na.rm = TRUE)
  })

  # Create network plot
  set.seed(42)
  ggraph(g, layout = 'fr') +
    geom_edge_link(aes(width = abs_weight, color = weight),
                   alpha = 0.7, show.legend = TRUE) +
    scale_edge_width(range = c(1, 8), name = "Correlation\nStrength") +
    scale_edge_color_gradient2(low = "#E74C3C", mid = "#F39C12", high = "#27AE60",
                               midpoint = 0.5, limits = c(0, 1),
                               name = "Correlation\nValue") +
    geom_node_point(aes(size = avg_score, fill = avg_score),
                    shape = 21, color = "white", stroke = 2) +
    scale_size(range = c(15, 35), name = "Average\nScore") +
    scale_fill_gradient(low = "#3498DB", high = "#E67E22",
                        limits = c(1, 5), name = "Average\nScore") +
    geom_node_text(aes(label = name), size = 5.5, fontface = "bold",
                   color = "white") +
    theme_graph(base_family = "sans") +
    theme(
      legend.position = "right",
      legend.title = element_text(size = 11, face = "bold"),
      legend.text = element_text(size = 9),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.margin = margin(20, 20, 20, 20)
    ) +
    labs(
      title = "Resilience Dimension Correlation Network"
    )
} else {
  cat("Insufficient correlation strength for network visualization.\n")
}
```

### Network Insights

```{r network-insights}
if(nrow(edges) > 0) {
  strongest_link <- edges %>% arrange(desc(abs_weight)) %>% slice(1)
  weakest_link <- edges %>% arrange(abs_weight) %>% slice(1)

  # Calculate centrality
  centrality <- data.frame(
    Dimension = V(g)$name,
    Degree = degree(g),
    AvgScore = V(g)$avg_score
  ) %>% arrange(desc(Degree))

  most_connected <- centrality$Dimension[1]
  least_connected <- centrality$Dimension[nrow(centrality)]
}
```

**Strongest Correlation:** `r if(nrow(edges) > 0) paste(strongest_link$from, "↔", strongest_link$to)` (r = `r if(nrow(edges) > 0) sprintf("%.3f", strongest_link$weight)`)

- These dimensions develop in tandem - strengthening one typically strengthens the other.

**Most Connected Dimension:** `r if(nrow(edges) > 0) most_connected`

- This dimension serves as a hub, correlating with multiple other capabilities.
- Investing here creates ripple effects across the resilience system.

**Least Connected:** `r if(nrow(edges) > 0) least_connected`

- This dimension operates more independently.
- Requires dedicated attention as improvements elsewhere may not naturally boost it.

\newpage

## Phase-Level Correlation Network

Examining correlations within specific phases reveals how dimensions interact differently across the supply chain.

```{r phase-networks, fig.width=14, fig.height=14}
# Function to create phase network
create_phase_network <- function(phase_data, phase_name, color) {
  cor_matrix <- cor(phase_data, use = "pairwise.complete.obs")

  threshold <- 0.25
  edges <- data.frame()
  for(i in 1:(nrow(cor_matrix)-1)) {
    for(j in (i+1):ncol(cor_matrix)) {
      if(abs(cor_matrix[i,j]) >= threshold) {
        edges <- rbind(edges, data.frame(
          from = rownames(cor_matrix)[i],
          to = colnames(cor_matrix)[j],
          weight = cor_matrix[i,j],
          abs_weight = abs(cor_matrix[i,j])
        ))
      }
    }
  }

  if(nrow(edges) == 0) return(NULL)

  g <- graph_from_data_frame(edges, directed = FALSE)
  V(g)$avg_score <- sapply(V(g)$name, function(dim) {
    mean(phase_data[[dim]], na.rm = TRUE)
  })

  return(g)
}

# Create phase-specific networks
upstream_data <- df %>%
  transmute(Redundancy = up__r, Collaboration = up__c, Flexibility = up__f,
            Visibility = up__v, Agility = up__a)

internal_data <- df %>%
  transmute(Redundancy = in__r, Collaboration = in__c, Flexibility = in__f,
            Visibility = in__v, Agility = in__a)

downstream_data <- df %>%
  transmute(Redundancy = do__r, Collaboration = do__c, Flexibility = do__f,
            Visibility = do__v, Agility = do__a)

g_up <- create_phase_network(upstream_data, "Upstream", "#3498DB")
g_in <- create_phase_network(internal_data, "Internal", "#E67E22")
g_do <- create_phase_network(downstream_data, "Downstream", "#9B59B6")

# Plot all three networks
par(mfrow = c(2, 2), mar = c(2, 2, 3, 2))

plot_phase_network <- function(g, title, color) {
  if(!is.null(g)) {
    set.seed(42)
    coords <- layout_with_fr(g)

    # Edge colors and widths
    edge_colors <- sapply(E(g)$weight, function(w) {
      if(w > 0.6) color
      else if(w > 0.4) adjustcolor(color, alpha.f = 0.7)
      else adjustcolor(color, alpha.f = 0.4)
    })
    edge_widths <- E(g)$abs_weight * 10

    # Node sizes based on avg score
    node_sizes <- rescale(V(g)$avg_score, to = c(15, 40))

    plot(g,
         layout = coords,
         vertex.size = node_sizes,
         vertex.color = color,
         vertex.frame.color = "white",
         vertex.frame.width = 3,
         vertex.label.color = "white",
         vertex.label.font = 2,
         vertex.label.cex = 1.2,
         edge.width = edge_widths,
         edge.color = edge_colors,
         edge.curved = 0.2,
         main = title,
         main.cex = 1.5,
         main.font = 2)
  }
}

plot_phase_network(g_up, "Upstream Phase", "#3498DB")
plot_phase_network(g_in, "Internal Phase", "#E67E22")
plot_phase_network(g_do, "Downstream Phase", "#9B59B6")

# Legend
par(mar = c(0, 0, 0, 0))
plot.new()
legend("center",
       legend = c("Strong Correlation (>0.6)", "Moderate Correlation (0.4-0.6)",
                 "Weak Correlation (0.25-0.4)", "Node size = Avg Score"),
       lty = c(1, 1, 1, NA),
       lwd = c(4, 3, 2, NA),
       pch = c(NA, NA, NA, 21),
       pt.cex = c(NA, NA, NA, 2),
       pt.bg = c(NA, NA, NA, "#95A5A6"),
       col = c("#2C3E50", adjustcolor("#2C3E50", alpha.f = 0.7),
              adjustcolor("#2C3E50", alpha.f = 0.4), NA),
       bty = "n",
       cex = 1.2,
       title = "Network Legend",
       title.font = 2)
```

\newpage

## Correlation Heatmap Matrix

A traditional correlation matrix complements the network view, providing exact correlation values.

```{r correlation-heatmap, fig.width=10, fig.height=10}
# Enhanced correlation plot
col_palette <- colorRampPalette(c("#E74C3C", "#F8F9FA", "#27AE60"))(200)

corrplot(cor_matrix,
         method = "circle",
         type = "upper",
         tl.col = "black",
         tl.srt = 45,
         tl.cex = 1.3,
         cl.cex = 1.1,
         col = col_palette,
         addCoef.col = "black",
         number.cex = 1.4,
         number.font = 2,
         diag = FALSE,
         title = "Dimension Correlation Matrix",
         mar = c(0, 0, 2, 0))
```

### Correlation Patterns

```{r correlation-patterns}
# Extract correlation insights
high_cors <- which(cor_matrix > 0.6 & cor_matrix < 1, arr.ind = TRUE)
low_cors <- which(cor_matrix < 0.3 & cor_matrix > -0.3, arr.ind = TRUE)
```

**High Correlations (>0.6):** `r nrow(high_cors)` dimension pairs show strong positive relationships.

- These represent synergistic capabilities that reinforce each other.
- Training and development programs should target these pairs together.

**Independent Dimensions (<0.3):** Dimensions with low correlations require separate strategic focus.

- Don't expect improvements in one to automatically boost the other.
- Balanced investment across uncorrelated dimensions ensures comprehensive coverage.

\newpage

## Strategic Implications

### Network-Based Strategy

1. **Hub Strategy:** Focus on the most connected dimension (`r if(nrow(edges) > 0) most_connected`) to create cascading improvements across multiple areas.

2. **Pair Programming:** Develop the strongest correlated pairs together through integrated initiatives rather than siloed programs.

3. **Independent Investment:** Dimensions with weak correlations need dedicated resources and cannot be improved indirectly.

4. **Phase Differences:** If phase networks differ significantly, it indicates inconsistent implementation approaches that may need standardization.

### Practical Applications

**For High Correlation Pairs:**
- Joint training programs
- Integrated KPIs
- Combined technology solutions
- Shared governance structures

**For Low Correlation Pairs:**
- Separate workstreams
- Distinct budgets
- Specialized expertise
- Independent measurement systems

### Benchmarking Guidance

```{r benchmark-guidance}
# Calculate overall network density
if(nrow(edges) > 0) {
  network_density <- edge_density(g)
  density_level <- ifelse(network_density > 0.7, "highly integrated",
                         ifelse(network_density > 0.5, "moderately integrated",
                               "loosely integrated"))
} else {
  density_level <- "insufficient data"
}
```

Your resilience dimensions are **`r if(nrow(edges) > 0) density_level`** (network density: `r if(nrow(edges) > 0) sprintf("%.2f", network_density)`).

- **Highly Integrated (>0.7):** Strong systemic approach; focus on maintaining connections.
- **Moderately Integrated (0.5-0.7):** Good foundation; opportunities to strengthen links.
- **Loosely Integrated (<0.5):** Siloed capabilities; consider more integrated strategies.

---

*Report generated with ResilienceScan - Supply Chain Resilience Assessment Tool*
