---
title: "Supply Chain Resilience Distribution Analysis"
subtitle: "Violin Plot Visualization of Score Patterns"
format:
  pdf:
    documentclass: article
    geometry: margin=0.75in
    fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(viridis)

# Load data - find the project root dynamically
script_dir <- if (exists("knitr")) dirname(knitr::current_input()) else getwd()
project_root <- if (basename(script_dir) == "report_variations") dirname(script_dir) else script_dir
data_path <- file.path(project_root, "data", "cleaned_master.csv")
df <- read.csv(data_path, stringsAsFactors = FALSE)
```

# Executive Summary

This violin plot analysis reveals the distribution patterns of resilience scores across dimensions and phases. Unlike simple averages, these visualizations show the full spectrum of performance - including variability, outliers, and concentration points.

## Survey Overview

```{r overview}
total_companies <- length(unique(df$company_name[df$company_name != "" & !is.na(df$company_name)]))
total_respondents <- nrow(df)

cat(sprintf("**Organizations Surveyed:** %d\n\n", total_companies))
cat(sprintf("**Total Responses:** %d\n\n", total_respondents))
cat("\n**Distribution Analysis:** Violin plots reveal not just average performance but the full range and concentration of resilience capabilities.\n")
```

\newpage

## Dimension Score Distribution

The width of each violin indicates the concentration of scores at that level. Wider sections show where most organizations cluster, while narrow sections indicate less common scores.

```{r dimension-violin, fig.width=13, fig.height=8}
# Prepare dimension data
dimension_dist <- df %>%
  select(up__r, up__c, up__f, up__v, up__a,
         in__r, in__c, in__f, in__v, in__a,
         do__r, do__c, do__f, do__v, do__a) %>%
  summarise(
    Redundancy = c(up__r, in__r, do__r),
    Collaboration = c(up__c, in__c, do__c),
    Flexibility = c(up__f, in__f, do__f),
    Visibility = c(up__v, in__v, do__v),
    Agility = c(up__a, in__a, do__a)
  ) %>%
  pivot_longer(everything(), names_to = "Dimension", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  mutate(Dimension = factor(Dimension,
                            levels = c("Redundancy", "Collaboration", "Flexibility",
                                     "Visibility", "Agility")))

ggplot(dimension_dist, aes(x = Dimension, y = Score, fill = Dimension)) +
  geom_violin(trim = FALSE, alpha = 0.7, scale = "width") +
  geom_boxplot(width = 0.15, fill = "white", alpha = 0.5, outlier.alpha = 0.3) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4,
               fill = "red", color = "darkred") +
  scale_fill_viridis_d(option = "plasma", begin = 0.2, end = 0.9) +
  scale_y_continuous(breaks = 1:5, limits = c(0.5, 5.5)) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 0, size = 12, face = "bold"),
    axis.text.y = element_text(size = 11),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  labs(
    y = "Resilience Score (1-5)",
    title = "Distribution of Resilience Scores by Dimension"
  ) +
  annotate("text", x = 5.5, y = 0.8, label = "◆ = Mean Score",
           hjust = 1, size = 4, color = "darkred", fontface = "bold")
```

### Distribution Insights

```{r dim-insights}
# Calculate distribution statistics
dim_stats <- dimension_dist %>%
  group_by(Dimension) %>%
  summarise(
    Mean = mean(Score, na.rm = TRUE),
    Median = median(Score, na.rm = TRUE),
    SD = sd(Score, na.rm = TRUE),
    IQR = IQR(Score, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean))

most_consistent <- dim_stats %>% filter(SD == min(SD)) %>% pull(Dimension)
most_variable <- dim_stats %>% filter(SD == max(SD)) %>% pull(Dimension)
```

- **Most Consistent:** `r most_consistent` shows the least variability (SD: `r sprintf("%.3f", min(dim_stats$SD))`), indicating uniform performance across organizations.

- **Most Variable:** `r most_variable` displays the highest variability (SD: `r sprintf("%.3f", max(dim_stats$SD))`), suggesting significant differences in how organizations approach this dimension.

\newpage

## Phase Distribution Comparison

Comparing distributions across operational phases reveals where resilience implementation is most consistent versus most fragmented.

```{r phase-violin, fig.width=13, fig.height=8}
# Phase distribution data
phase_dist <- df %>%
  select(up__r, up__c, up__f, up__v, up__a,
         in__r, in__c, in__f, in__v, in__a,
         do__r, do__c, do__f, do__v, do__a) %>%
  summarise(
    Upstream = c(up__r, up__c, up__f, up__v, up__a),
    Internal = c(in__r, in__c, in__f, in__v, in__a),
    Downstream = c(do__r, do__c, do__f, do__v, do__a)
  ) %>%
  pivot_longer(everything(), names_to = "Phase", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  mutate(Phase = factor(Phase, levels = c("Upstream", "Internal", "Downstream")))

ggplot(phase_dist, aes(x = Phase, y = Score, fill = Phase)) +
  geom_violin(trim = FALSE, alpha = 0.7, scale = "width", size = 1) +
  geom_boxplot(width = 0.2, fill = "white", alpha = 0.6, outlier.size = 2) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 5,
               fill = "red", color = "darkred") +
  scale_fill_manual(values = c("Upstream" = "#3498DB",
                                "Internal" = "#E67E22",
                                "Downstream" = "#9B59B6")) +
  scale_y_continuous(breaks = 1:5, limits = c(0.5, 5.5)) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(angle = 0, size = 13, face = "bold"),
    axis.text.y = element_text(size = 11),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  labs(
    y = "Resilience Score (1-5)",
    title = "Score Distribution Across Supply Chain Phases"
  )
```

### Phase Pattern Analysis

```{r phase-patterns}
phase_stats <- phase_dist %>%
  group_by(Phase) %>%
  summarise(
    Mean = mean(Score, na.rm = TRUE),
    SD = sd(Score, na.rm = TRUE),
    Q25 = quantile(Score, 0.25, na.rm = TRUE),
    Q75 = quantile(Score, 0.75, na.rm = TRUE),
    .groups = "drop"
  )

most_uniform_phase <- phase_stats %>% filter(SD == min(SD)) %>% pull(Phase)
most_diverse_phase <- phase_stats %>% filter(SD == max(SD)) %>% pull(Phase)
```

- **Most Uniform Phase:** `r most_uniform_phase` (SD: `r sprintf("%.3f", min(phase_stats$SD))`) - Organizations show similar capabilities in this phase.

- **Most Diverse Phase:** `r most_diverse_phase` (SD: `r sprintf("%.3f", max(phase_stats$SD))`) - Wide variation in capabilities suggests best practice gaps.

\newpage

## Detailed Phase-Dimension Distribution Matrix

This comprehensive view shows the distribution for each specific dimension within each phase, revealing granular patterns.

```{r matrix-violin, fig.width=14, fig.height=11}
# Detailed phase-dimension data
detailed_dist <- df %>%
  select(up__r, up__c, up__f, up__v, up__a,
         in__r, in__c, in__f, in__v, in__a,
         do__r, do__c, do__f, do__v, do__a) %>%
  pivot_longer(everything(), names_to = "Metric", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  separate(Metric, into = c("Phase", "Dimension"), sep = "__") %>%
  mutate(
    Phase = case_when(
      Phase == "up" ~ "Upstream",
      Phase == "in" ~ "Internal",
      Phase == "do" ~ "Downstream"
    ),
    Dimension = case_when(
      Dimension == "r" ~ "Redundancy",
      Dimension == "c" ~ "Collaboration",
      Dimension == "f" ~ "Flexibility",
      Dimension == "v" ~ "Visibility",
      Dimension == "a" ~ "Agility"
    ),
    Phase = factor(Phase, levels = c("Upstream", "Internal", "Downstream")),
    Dimension = factor(Dimension, levels = c("Redundancy", "Collaboration",
                                             "Flexibility", "Visibility", "Agility"))
  )

ggplot(detailed_dist, aes(x = Dimension, y = Score, fill = Phase)) +
  geom_violin(trim = FALSE, alpha = 0.6, scale = "width", position = position_dodge(0.9)) +
  geom_boxplot(width = 0.15, position = position_dodge(0.9),
               fill = "white", alpha = 0.5, outlier.size = 0.5) +
  scale_fill_manual(values = c("Upstream" = "#3498DB",
                                "Internal" = "#E67E22",
                                "Downstream" = "#9B59B6")) +
  scale_y_continuous(breaks = 1:5, limits = c(0.5, 5.5)) +
  facet_wrap(~Phase, ncol = 1) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 13, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y = element_text(size = 10),
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold"),
    strip.background = element_rect(fill = "gray90", color = "gray50"),
    panel.grid.major.x = element_blank(),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
  ) +
  labs(
    y = "Resilience Score (1-5)",
    title = "Comprehensive Phase-Dimension Distribution Analysis"
  )
```

\newpage

## Company Size Distribution Patterns

Examining how score distributions vary by company size reveals whether larger organizations truly have more consistent resilience.

```{r size-violin, fig.width=13, fig.height=8}
# Size-based analysis
size_dist <- df %>%
  filter(!is.na(size_number_of_employees) & size_number_of_employees != "") %>%
  mutate(
    Size_Category = case_when(
      grepl("1-9|10-49", size_number_of_employees) ~ "Small\n(1-49)",
      grepl("50-99|100-499", size_number_of_employees) ~ "Medium\n(50-499)",
      grepl("500-999|1,000-4,999|5,000-9,999", size_number_of_employees) ~ "Large\n(500-9,999)",
      grepl("10,000", size_number_of_employees) ~ "Enterprise\n(10,000+)",
      TRUE ~ NA_character_
    ),
    Overall = overall_scres
  ) %>%
  filter(!is.na(Size_Category) & !is.na(Overall)) %>%
  mutate(Size_Category = factor(Size_Category,
                                levels = c("Small\n(1-49)", "Medium\n(50-499)",
                                          "Large\n(500-9,999)", "Enterprise\n(10,000+)")))

if (nrow(size_dist) > 0) {
  ggplot(size_dist, aes(x = Size_Category, y = Overall, fill = Size_Category)) +
    geom_violin(trim = FALSE, alpha = 0.7, scale = "width") +
    geom_boxplot(width = 0.15, fill = "white", alpha = 0.5, outlier.alpha = 0.5) +
    stat_summary(fun = mean, geom = "point", shape = 23, size = 4,
                 fill = "red", color = "darkred") +
    scale_fill_brewer(palette = "Set2") +
    scale_y_continuous(breaks = 1:5, limits = c(1, 5)) +
    theme_minimal() +
    theme(
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 14, face = "bold"),
      axis.text.x = element_text(angle = 0, size = 11, face = "bold"),
      axis.text.y = element_text(size = 11),
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)
    ) +
    labs(
      y = "Overall Resilience Score",
      title = "Resilience Score Distribution by Company Size"
    )
} else {
  cat("Insufficient size data for distribution analysis.\n")
}
```

\newpage

## Key Findings & Strategic Implications

### Distribution Patterns Tell a Story

1. **Bimodal Distributions**: If violin plots show two bulges, it indicates two distinct groups - high and low performers - suggesting best practices aren't widely adopted.

2. **Narrow Distributions**: Violin shapes that are consistently narrow indicate mature, standardized practices across organizations.

3. **Wide Distributions**: Broad, spread-out violins reveal significant variability and opportunity for learning from top performers.

4. **Outliers**: Points outside the violin shapes represent exceptional cases worthy of detailed investigation.

### Actionable Insights

```{r actionable}
# Overall variability assessment
overall_sd <- sd(df$overall_scres, na.rm = TRUE)
variability_level <- ifelse(overall_sd < 0.5, "low",
                           ifelse(overall_sd < 0.8, "moderate", "high"))
```

**Overall Variability:** `r variability_level` (SD: `r sprintf("%.3f", overall_sd)`)

**Recommendations:**

1. **High Variability Dimensions**: Focus benchmarking and best practice sharing on `r most_variable` where organizations show the widest performance gaps.

2. **Consistent Dimensions**: For `r most_consistent`, shift from broad improvement programs to targeted excellence initiatives for laggards.

3. **Phase Strategy**: The `r most_diverse_phase` phase shows the most opportunity for improvement through standardization and capability building.

4. **Size Considerations**: Analyze whether your size category's distribution matches expectations - smaller firms may have wider distributions due to resource constraints.

5. **Aspire to the Upper Quartile**: Organizations should benchmark against the 75th percentile of their violin plot, not just the mean, to set ambitious but achievable targets.

---

*Report generated with ResilienceScan - Supply Chain Resilience Assessment Tool*
